Numerical modelling of the Cordilleran ice sheet

### 2014-08-10 kappa.tex
Implement Irina's corrections on abstract.

This doctoral dissertation presents a study of the glacial history of the North
American Cordillera using numerical ice sheet modelling calibrated against
field evidence. This area, characterized by the steep topography of several
mountain ranges separated by large inter-montane depressions, was once covered
by a large-scale ice mass: the former Cordilleran ice sheet. Because of the
irregular topography on which the ice sheet formed, geological studies have
often only had local or regional relevance, and thus the Cordilleran ice sheet
remains the least understood Pleistocene ice sheet in terms of its extent,
volume, and dynamics.

In this thesis, I present numerical simulations performed with the Parallel Ice
Sheet Model (PISM), that allow quantitative reconstructions of the former ice
sheet evolution based on approximated physics of glacier flow. These
simulations show that the geometry of the Last Glacial Maximum Cordilleran ice
sheet was largely controlled by sharp contrasts in regional temperature,
precipitation, and daily temperature variability associated with the presence
of the mountain ranges.

However, this maximum stage appear short-lived and out of balance with
contemporaneous climate. During most of the simulated last glacial cycle, the
North American Cordillera is characterized by an intermediate state of
glaciation including isolated glaciers and ice caps covering the major mountain
ranges, the largest of which is located over the Skeena Mountains. Numerically
modelled patterns of ice sheet growth and decay are diverse, and reveal a
higher complexity than that encapsulated in previous, conceptual models.

The simulations presented in this thesis demonstrate the potential of numerical
ice sheet modelling to inform on ice sheet history over glacial cycles,
provided that ice sheet models can be calibrated against field constraints.

>>>>

This doctoral dissertation presents a study of the glacial history of the North
American Cordillera using numerical ice sheet modelling calibrated against
field evidence. This area, characterized by the steep topography of several
mountain ranges separated by large inter-montane depressions, was once covered
by a large-scale ice mass: the former Cordilleran ice sheet. Because of the
irregular topography on which the ice sheet formed, geological studies have
often had only local or regional relevance, thus leaving the Cordilleran ice
sheet least understood among Pleistocene ice sheets in terms of its extent,
volume, and dynamics.

Here, I present numerical simulations that allow quantitative reconstructions
of the former ice
sheet evolution based on approximated physics of glacier flow. These
simulations show that the geometry of the Last Glacial Maximum Cordilleran ice
sheet was largely controlled by sharp contrasts in regional temperature,
precipitation, and daily temperature variability associated with the presence
of mountain ranges.

However, this maximum stage appears short-lived and out of balance with
contemporaneous climate. During most of the simulated last glacial cycle, the
North American Cordillera is characterized by an intermediate state of
glaciation including isolated glaciers and ice caps covering major mountain
ranges, the largest of which is located over the Skeena Mountains. The numerically modelled Cordilleran ice sheet appears in constant imbalance with evolving climate conditions, while the nature of this transient response transcends the complexity of two-dimensional, conceptual models of ice sheet growth and decay.

This thesis demonstrates the potential of numerical ice sheet modelling to
inform on ice sheet history and former climate conditions over a glacial cycle,
given that ice sheet models can be calibrated against field constraints.
